---
title: "DDI analysis over signor network "
author: "Livia Perfetto"
date: "08/08/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


#### Libraries
```{r message=FALSE,warning=FALSE}
library(RCy3)
library(igraph)
library(data.table)
library(httr) 
library(tidyr)
library(dplyr)
library(readxl)
library(tidyverse)
library(xlsx)
library(org.Hs.eg.db)
library(AnnotationDbi)
```

### define script functions
```{r bins.per.pair}

`%notin%` <- Negate(`%in%`) 


#retrieve_edges_connect_add_bridge takes as input signor dataframe and a list of input proteins and retrieve all the edges in the connect + first neigh mode
retrieve_edges_connect_add_bridge <- function(input_proteins){
 
   ## the function perform a query in SIGNOR using API and connection method level 2 (First neighbours)
  signor_FN_query <-
  paste0('https://signor.uniroma2.it/getData.php?type=connect&proteins=',
         paste0(input_proteins, collapse = '%2C'),
                '&level=2')
  
  signor_FN <- as.data.frame(fread(signor_FN_query))
  
  ##for the network it retrieves a list of edge names (defined as regulator -> target)
 
  signor_FN$edge_key <- paste0 (signor_FN$V1, '|', signor_FN$V5)
  
  edge_names <- list(signor_FN$edge_key )
  
  return(edge_names)
  
}


retrieve_nodes_connect_add_bridge <- function(input_proteins){
  
  ## the function perform a query in SIGNOR using API and connection method level 2 (First neighbours)
  signor_FN_query <-
  paste0('https://signor.uniroma2.it/getData.php?type=connect&proteins=',
         paste0(input_proteins, collapse = '%2C'),
                '&level=2')
  
  signor_FN <- as.data.frame(fread(signor_FN_query))
  
  signor_FN$edge_key <- paste0 (signor_FN$V1, '|', signor_FN$V5)
  
  ##for the network it retrieves a list of node names
  nodes_names <- list(unique(signor_FN$V1, signor_FN$V5))
  
  return(nodes_names)
  
}


##create_matrix function takes a list of lists and generate a matrix of intersections

create_matrix <- function(l){sapply(seq_len(length(l)), function(x) ## where l = list of lists
  
  sapply(seq_len(length(l)), function(y) length(intersect(unlist(l[x]), unlist(l[y])))))}


```

### import SIGNOR data

```{r bins.per.pair, echo = FALSE,warning=FALSE}
###get network from signor

# Signor Relation reading
signor = as.data.frame(fread(paste0('https://signor.uniroma2.it/getDataInternal.php?&localization&organism=9606'), header=TRUE))


```

###START ANALYSIS IN THE CONTEXT OF NEURO


### extract datasets to analyse, NDD as defined by PMCI and ASC
```{r bins.per.pair}

##import NDD genes curated by the PCMI (Willsey et al, 2018, https://www.cell.com/cell/pdf/S0092-8674(18)30784-0.pdf)

read_xlsx('input_tables/all_NDD.xlsx', 
          sheet = 1, range = NULL, 
          col_names = TRUE,
          col_types = NULL, 
          na = "", 
          trim_ws = TRUE, 
          skip = 0,
          progress = readxl_progress(), .name_repair = "unique")-> NDD


NDD%>%
  filter(Disease == 'ASD') -> ASD
NDD%>%
  filter(Disease == 'EE') -> EE
NDD%>%
  filter(Disease == 'ID') -> ID
NDD%>%
  filter(Disease == 'SCZ') -> SCZ

## extract independent ASC list of ASD genes

ASC_file = 'input_tables/ASC_gene_results'

read.csv(ASC_file, header = T, sep = '\t') -> ASC

ASC %>%
  filter(grepl('ENS', gene_id) & qval < 0.05) -> ASC

ASC$SYMBOL <- AnnotationDbi::mapIds(org.Hs.eg.db, 
                      keys = ASC$gene_id,
                      keytype = 'ENSEMBL',
                      column = 'SYMBOL')

```

## define names of the comparison
```{r bins.per.pair, echo = FALSE,warning=FALSE}

PCMI_names <- c('PCMI-ASD','ASC-ASD', 'PCMI-SCZ','PCMI-EE','PCMI-ID')
# ASC_names <- c('ASC-ASD','PCMI-SCZ','PCMI-EE','PCMI-ID')

```


### create subnetworks of SIGNOR, for specific lists
## ASD genes source: PCMI
## edge type: common genes
```{r bins.per.pair}

setwd('~/Desktop/neuro/PAPER/analysis_resubmission/DDI_analysis/')

ASD_list_PCMI <- list(ASD$Gene)
ASD_list_ASC <- list(ASC$SYMBOL)
SCZ_list_PCMI <- list(SCZ$Gene)
ID_list_PCMI <- list(ID$Gene)
EE_list_PCMI <- list(EE$Gene)

list_of_lists <- list(ASD_list_PCMI,
                      ASD_list_ASC,
                      SCZ_list_PCMI,
                      EE_list_PCMI,
                      ID_list_PCMI)


create_matrix (list_of_lists) -> matrix_of_intersection_genes_ASD_PCMI

row.names(matrix_of_intersection_genes_ASD_PCMI) <- PCMI_names
colnames(matrix_of_intersection_genes_ASD_PCMI) <- PCMI_names
matrix_of_intersection_genes_ASD_PCMI

## create a DDI from the comparison matrix 
graph_from_adjacency_matrix(matrix_of_intersection_genes_ASD_PCMI,
                            mode = "undirected",
                            weighted = T) -> graph_genes_ASD_PCMI

graph_genes_ASD_PCMI <- igraph::simplify(graph_genes_ASD_PCMI, remove.loops = T)

createNetworkFromIgraph(
  graph_genes_ASD_PCMI,
  title = paste0("ASD genes source: ALL, edge type: shared genes"),
  collection = "DDI analysis")

```

### create subnetworks of SIGNOR, for specific lists
## ASD genes source: PCMI
## edge type: common edges
```{r bins.per.pair}

ASD_list_PCMI_edge <- retrieve_edges_connect_add_bridge(ASD$Gene)
ASD_list_ASC_edge <- retrieve_edges_connect_add_bridge(ASC$SYMBOL)

# SCZ_list_PCMI_edge  <- retrieve_edges_connect_add_bridge(SCZ$Gene)
SCZ_list_PCMI_edge <- list()
ID_list_PCMI_edge  <- retrieve_edges_connect_add_bridge(ID$Gene)
EE_list_PCMI_edge  <- retrieve_edges_connect_add_bridge(EE$Gene)

list_of_lists <- list(ASD_list_PCMI_edge ,
                      ASD_list_ASC_edge ,
                      SCZ_list_PCMI_edge ,
                      EE_list_PCMI_edge ,
                      ID_list_PCMI_edge )


create_matrix (list_of_lists) -> matrix_of_intersection_edges_ASD_PCMI

row.names(matrix_of_intersection_edges_ASD_PCMI) <- PCMI_names
colnames(matrix_of_intersection_edges_ASD_PCMI) <- PCMI_names
matrix_of_intersection_edges_ASD_PCMI

## create a DDI from the comparison matrix 
graph_from_adjacency_matrix(matrix_of_intersection_edges_ASD_PCMI,
                            mode = "undirected",
                            weighted = T) -> graph_genes_ASD_PCMI

graph_genes_ASD_PCMI <- igraph::simplify(graph_genes_ASD_PCMI, remove.loops = T)

createNetworkFromIgraph(
  graph_genes_ASD_PCMI,
  title = paste0("ASD genes source: ALL, edge type: shared edges"),
  collection = "DDI analysis")
```
### create subnetworks of SIGNOR, for specific lists
## ASD genes source: PCMI
## edge type: common edges
```{r bins.per.pair}

ASD_list_PCMI_node <- retrieve_nodes_connect_add_bridge(ASD$Gene)
ASD_list_ASC_node  <- retrieve_nodes_connect_add_bridge(ASC$SYMBOL)

#SCZ_list_PCMI_node <- retrieve_nodes_connect_add_bridge(SCZ$Gene)
SCZ_list_PCMI_node  <- list()
ID_list_PCMI_node  <- retrieve_nodes_connect_add_bridge(ID$Gene)
EE_list_PCMI_node  <- retrieve_nodes_connect_add_bridge(EE$Gene)

list_of_lists <- list(ASD_list_PCMI_node ,
                      ASD_list_ASC_node,
                      SCZ_list_PCMI_node ,
                      EE_list_PCMI_node,
                      ID_list_PCMI_node)


create_matrix (list_of_lists) -> matrix_of_intersection_nodes_ASD_PCMI

row.names(matrix_of_intersection_nodes_ASD_PCMI) <- PCMI_names
colnames(matrix_of_intersection_nodes_ASD_PCMI) <- PCMI_names
matrix_of_intersection_nodes_ASD_PCMI

## create a DDI from the comparison matrix 
graph_from_adjacency_matrix(matrix_of_intersection_nodes_ASD_PCMI,
                            mode = "undirected",
                            weighted = T) -> graph_nodes_ASD_PCMI

graph_nodes_ASD_PCMI <- igraph::simplify(graph_nodes_ASD_PCMI, remove.loops = T)

createNetworkFromIgraph(
  graph_nodes_ASD_PCMI,
  title = paste0("ASD genes source: ALL, edge type: shared nodes"),
  collection = "DDI analysis")
```

```